\textbf{ \Large {Decentralized autonomous sensor fault detection using neural networks}}

### 2015-02-16 Paper/PaperJahrSchlich.tex
new dataflow figure, notes to labex
letzte Zwischenabgabe

\documentclass[12pt,a4paper]{scrartcl}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[color=yellow!15]{todonotes}
\usepackage{graphicx}
\usepackage[numbers, sort&compress, square]{natbib}
\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{titlesec}
\usepackage{color}
\usepackage{booktabs}

@@

\makeatletter
\renewcommand\section{\@startsection{section}{1}{\z@}%
                     {-7.5ex \@plus -1ex \@minus -10.2ex}%
                     {2.3ex \@plus.2ex}%
                     {\sffamily\large\bfseries}}
\makeatother

@@

Traditionally, the sensor nodes are connected to computer systems with cables.
Using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.

@@

In this study, analytical redundancy is implemented into wireless sensor nodes based on artificial neural networks.
Artificial neural networks essentially consist of interconnected data processing units called neurons. 
The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.

This paper is organized as follows:
First, a wireless structural health monitoring system is designed and implemented. 
Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
Then, the system is tested in laboratory experiments. 
Finally, the experimental results are discussed and future research directions are proposed.

@@

\newpage

@@

The SHM system performs the following tasks:
1. data acquisition,
2. data processing,
3. data transmission, 
4. data storage,
5. diagnostics and 
6. information retrieval.
The dataflow is illustrated in \autoref{fig:flow}.
Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a FFT to determine the characteristics of the oscillation of the structure. 
The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
On the host computer, tasks 4 is conducted by storing the data in a MySQL database.
Task 5 and 6---additional analysis and diagnosis---are conducted on the host computer in further steps.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/dataflow_tasks_nn.eps}
    \caption{Dataflow and associated SHM tasks}
    \label{fig:flow}
\end{figure}

\colorbox{cyan}{section ab hier teilweise neu (NN):}
The SHM system is programmed object-oriented in Java. 
Object orientation uses objects, which are instanciated using classes as paradigm. 
A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
A package organizes several Java classes that build a program.

\autoref{fig:UML-sn} describes the classes of the \texttt{sensor\-node} package.
The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Measure\-ment}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the Java library "SNIPE" \todo{Quelle für SNIPE!} that implement a framework for neural networks.
The classes are embedded directly into the sensor nodes.

The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a FFT on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.

The \texttt{Measure\-ment} class defines analyzed measurements by storing the address of the sensor node, the maximal magnitude, the natural frequency and if a sensor defect was detected.
For every measurement, one object of the type \texttt{Measure\-ment} is instantiated.
Using the SNIPE library, a neural network is defined and created for fault detection.
The \texttt{Network\-Training} class defines a routine to train the neural network with prior measured \texttt{Measure\-ment} objects.
The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer the \texttt{Measure\-ment} objects from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE \texttt{Neural\-Network} class, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the measurement.

\begin{figure}[htb]
    \centering
    \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
    \caption{Class diagram of the \texttt{sensornode} package}
    \label{fig:UML-sn}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
    \caption{Class diagram of the \texttt{basestation} package}
    \label{fig:UML-bs}
\end{figure}

\section*{Implementation and training of neural networks}

Neural networks are implemented into the sensor nodes using SNIPE, a open source Java Library.
\begin{figure}[ht]
    \centering
    \includegraphics{figures/neuralnetwork.pdf}
    \caption{Schematic drawing of an artificial neural network with three layers}
    \label{fig:neuralnetwork}
\end{figure}
\todo[inline]{Grafik wird an unser neuronales Netz angepasst}

%\begin{itemize}
%\item neurons
%\item layer
%\item weights
%\item activation function / identity function
%\end{itemize}

%\textbf{2. paragraph: proposed nn}
%
%\begin{itemize}
%\item SNIPE? implementation in java?
%\item activation function / identity function
%\end{itemize}
%
%\textbf{3. paragraph: learning general}
%
%\begin{itemize}
%\item set with test values
%\item iterative weight adjustment
%\end{itemize}
%
%\textbf{4. paragraph?: learning, proposed}

@@

In the following section, the laboratory experiments are described.
First, a description of the test setup is given, second, the data aquisition and processing are depicted, and finally, the results are discussed.

@@

The structure is excited by deflecting and releasing the top of the structure.
This excitation method ensures a free vibration in natural frequency with little interferences.
After excitation, when the acceleration threshold is exceeded, the sensor nodes automatically start measuring the acceleration.
To minimize the wireless data traffic, each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building \cite{rao2011fast}.

\textit{The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
The natural frequencies can be identified by the peaks in the magnitude graph.
The first two natural frequencies are located at approximately 2.3\,Hz (1\textsuperscript{st} natural frequency) and 8\,Hz (2\textsuperscript{nd} natural frequency).
Each sensor node transfers the values of the first natural frequency and the corresponding magnitude to the basestation. 
The values are stored in a MySQL database and analyzed.}

\begin{figure}[htb]
    \centering
    \includegraphics[scale=0.9]{figures/76HzHalf.pdf}
    \caption{frequency domain graph of the oscillation}
    \label{fig:76hzfft}
\end{figure}

\textit{The natural frequencies have been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values---and the results of a FE analysis.
The highest possible sampling rate was determined at 76\,Hz.
Higher samping rates could not be handled by the sensor nodes.
Test runs with more than 512 measured values showed no significant increase in precision.
Therefore, a sampling rate of 76\,Hz and 512 measured values were chosen.}

\begin{itemize}
\item neuronales Netz
\item TRAININGSPHASE
\item Austausch der gemessenen FFT zwischen den Sensoren
\item trainieren des neuronalen Netzes, bis keine Verbesserung bei Vorhersagegenauigkeit eintritt
\item MESSPHASE
\item Austausch der gemessenen FFT zwischen den Sensoren
\item Nutzen des neuronalen Netzes, um eventuellen Fehler zu finden
\item KÜNSTLICHEN FEHLER ERZEUGEN
\item TABELLE MIT ERMITTELTEN / GEMESSENEN WERTEN; MIT / OHNE FEHLER
\end{itemize}

%\textit{Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
%Example results are shown in \autoref{tab:lab-ex}.
%The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.}
%
%\textit{Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
%A number stored in the used data format double occupies 8\,bytes of space.
%The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.}

\begin{table}[hb]
	\centering
	\begin{tabular}{l l c c c}
		\toprule
		ID & Position & Magnitude & Natural frequency [Hz] & Sampling rate [Hz]\\ 
		\midrule
		1 & top & 3.78 & 2.2461 & 76\\ 
		1 & middle & 3.23 & 2.2461 & 76\\ 
		1 & bottom & 1.97 & 2.2461 & 76\\ 
		\midrule	
		2 & top & 5.74 & 2.2536 & 76\\  
		2 & middle & 4.38 & 2.2536 & 76\\ 
		2 & bottom & 2.55 & 2.2536 & 76\\
		\midrule
		3 & top & 8.20 & 2.2536 & 76\\
		3 & middle & 6.35 & 2.2536 & 76\\ 
		3 & bottom & 3.79 & 2.2536 & 76\\
		\bottomrule
	\end{tabular}
	\caption{Natural frequencies of the test structure}
	\label{tab:lab-ex}
\end{table}

>>>>

\documentclass[12pt,a4paper]{scrartcl}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[color=yellow!15]{todonotes}
\usepackage{graphicx}
\usepackage[numbers, sort&compress, square]{natbib}

\usepackage{titlesec}
\usepackage{color}
\usepackage{booktabs}

@@

\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}

%\makeatletter
%\renewcommand\section{\@startsection{section}{1}{\z@}%
%                     {-7.5ex \@plus -1ex \@minus -10.2ex}%
%                     {2.3ex \@plus.2ex}%
%                     {\sffamily\large\bfseries}}
%\makeatother

@@

Traditionally, the sensor nodes are connected to computer systems with cables.
Using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems. \colorbox{cyan}{hinzugefügt: }However, when using wireless SHM systems, special attention has to be paid to data transfer. 
To save transmission time and storage space, and to decrease the probability of data loss, the transmitted data is reduced using embedded computing.

@@

In this study, analytical redundancy is implemented into wireless sensor nodes based on artificial neural networks.
Artificial neural networks essentially consist of interconnected data processing units called neurons. 
The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.

This paper is organized as follows:
First, a dezentralized wireless structural health monitoring system is designed and implemented. 
Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
Then, the system is tested in laboratory experiments. 
Finally, the experimental results are discussed and future research directions are proposed.

@@

@@

The SHM system performs the following tasks:
1. data acquisition,
2. data processing,
3. data transmission, 
4. data storage,
5. diagnostics and 
6. information retrieval, see \citet{BisbySHM}.
The dataflow is illustrated in \autoref{fig:flow}.
Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a FFT to determine the characteristics of the oscillation of the structure.
The results of the FFT are fed into a neural network to detect sensor faults.
The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
On the host computer, tasks 4 is conducted by storing the data in a MySQL database.
Task 5 and 6---additional analysis and diagnosis---are conducted on the host computer in further steps.

\begin{figure}[hbtp]
    \centering
    \includegraphics{figures/dataflow_tasks_nn.eps}
    \caption{Dataflow and associated SHM tasks}
    \label{fig:flow}
\end{figure}

The neural network represents the structure of the SHM system.
Each sensor node of the SHM system is represented by either an input neuron or an output neuron, as shown in \autoref{fig:neuralnetwork}.
During computation, each sensor node embeds a distinct neural network. 
Measurements of other sensor nodes are fed into the input neurons of the neural network; the output neuron returns the predicted measurement of the embedding sensor node. 
Hidden neurons are used for computation and do not represent a physical counterpart.

\begin{figure}[hbtp]
    \centering
    \includegraphics{figures/neuralnetwork.pdf}
    \caption{Scheme of the neural network}
    \label{fig:neuralnetwork}
\end{figure}

The SHM system is programmed object-oriented in Java. 
Object orientation uses objects, which are instanciated using classes as paradigm. 
A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
A package organizes several Java classes that build a program.

\autoref{fig:UML-sn} describes the classes of the \texttt{sensor\-node} package.
The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Measure\-ment}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the Java library "SNIPE" \citep{SNIPE} that implement a framework for neural networks.
The classes are embedded directly into the sensor nodes.

The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a FFT on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
The \texttt{Measure\-ment} class defines analyzed measurements by storing the address of the sensor node, the maximal magnitude, the natural frequency and if a sensor fault was detected during runtime.
For every measurement, one object of the type \texttt{Measure\-ment} is instantiated.
Using the SNIPE library, a neural network is defined and created for fault detection.
The \texttt{Network\-Training} class defines a routine to train the neural network with prior measured \texttt{Measure\-ment} objects.
The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer the \texttt{Measure\-ment} objects from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the measurement.

\begin{figure}[htbp]
    \centering
    \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
    \caption{Class diagram of the \texttt{sensornode} package ((Update auf NN folgt))}
    \label{fig:UML-sn}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
    \caption{Class diagram of the \texttt{basestation} package}
    \label{fig:UML-bs}
\end{figure}

@@

In the following section, the laboratory experiments are described.
First, a description of the test setup is given, second, the data aquisition, data processing and sensor fault detection are depicted, and, finally, the results are discussed.

@@

The structure is excited by deflecting and releasing the top of the structure.
This excitation method ensures a free vibration in natural frequency with little interferences.
After excitation, when the acceleration threshold is exceeded, the sensor nodes automatically start measuring the acceleration.
Each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building \cite{rao2011fast}.

\begin{itemize}
\item \textbf{neuronales Netz}
\item TRAININGSPHASE
\item Austausch der gemessenen FFT zwischen den Sensoren
\item trainieren des neuronalen Netzes, bis keine Verbesserung bei Vorhersagegenauigkeit eintritt
\item MESSPHASE
\item Austausch der gemessenen FFT zwischen den Sensoren
\item Nutzen des neuronalen Netzes, um eventuellen Fehler zu finden
\item KÜNSTLICHEN FEHLER ERZEUGEN
\item TABELLE MIT ERMITTELTEN / GEMESSENEN WERTEN; MIT / OHNE FEHLERN
\end{itemize}

%The neural network has been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values.

%The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
%The natural frequencies can be identified by the peaks in the magnitude graph.
%Each sensor node stores the calculated first natural frequency and the corresponding magnitude in a \texttt{Measurement} object and transmits the object to the other sensor nodes.

%\begin{figure}[htb]
%    \centering
%    \includegraphics{figures/76HzHalf.pdf}
%    \caption{frequency domain graph of the oscillation}
%    \label{fig:76hzfft}
%\end{figure}

After the sensor fault detection, the \texttt{Measurement} objects of all sensor nodes ar transmitted to the basestation.
The values and detected sensor faults are stored in a MySQL database and availeble for further analysis.

%Example results are shown in \autoref{tab:lab-ex}.
%The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.
%
%\textit{Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
%A number stored in the used data format double occupies 8\,bytes of space.
%The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.}

%\begin{table}[hb]
%	\centering
%	\begin{tabular}{l l c c c}
%		\toprule
%		ID & Position & Magnitude & Natural frequency [Hz] & Sampling rate [Hz]\\ 
%		\midrule
%		1 & top & 3.78 & 2.2461 & 76\\ 
%		1 & middle & 3.23 & 2.2461 & 76\\ 
%		1 & bottom & 1.97 & 2.2461 & 76\\ 
%		\midrule	
%		2 & top & 5.74 & 2.2536 & 76\\  
%		2 & middle & 4.38 & 2.2536 & 76\\ 
%		2 & bottom & 2.55 & 2.2536 & 76\\
%		\midrule
%		3 & top & 8.20 & 2.2536 & 76\\
%		3 & middle & 6.35 & 2.2536 & 76\\ 
%		3 & bottom & 3.79 & 2.2536 & 76\\
%		\bottomrule
%	\end{tabular}
%	\caption{Natural frequencies of the test structure}
%	\label{tab:lab-ex}
%\end{table}
